
% ================================================================
% CHAPTER 1: INTRODUCTION  (humanised rewrite)
% ================================================================

\section{Outline}

\subsection{General System}

Imagine you ask an AI assistant: ``Who is the current Prime Minister of India?''
The AI confidently says ``Narendra Modi.'' But your document collection says
something different---a newer document says a different leader has taken over.
Who should the system believe? This is exactly the problem that TruthfulRAG v5
tries to solve.

TruthfulRAG v5 is a Knowledge-Graph based question-answering pipeline. It reads
a collection of documents, builds a structured map of the facts in those
documents (a Knowledge Graph), and then---when a user asks a question---it does
not just search for keywords. It reasons over the graph, detects if two facts
contradict each other, figures out which one is more trustworthy based on how
recent it is and how many documents agree with it, and then gives a final answer
along with a score saying how confident it is.

The whole system runs on your own machine. There is no cloud, no subscription,
and no data leaving your computer. The LLM used is Qwen2.5-7B-Instruct, served
locally via Ollama. The graph database is Neo4j Community Edition.

\subsection{Overall Project Flow}

The system runs in four stages, one after the other:

\begin{enumerate}
    \item \textbf{Schema Inference} --- Before the system even starts building
    the graph, it reads a few sample documents and asks the LLM: ``What kinds
    of things are in these documents?'' The LLM figures out that the documents
    are about, say, Indian scientists, and returns categories like
    \texttt{PERSON}, \texttt{ORGANIZATION}, and \texttt{MISSION}. No manual
    setup is needed.

    \item \textbf{Graph Construction (Module A)} --- The system reads all the
    documents, pulls out facts (e.g., ``ISRO launched Chandrayaan-3 in 2023''),
    and stores them as nodes and arrows in a Neo4j database. It also keeps track
    of \emph{which year} each fact is from, and \emph{how many documents} said
    the same thing.

    \item \textbf{Graph Retrieval (Module B)} --- When the user asks a question,
    the system searches the graph. It does not just look for matching words; it
    uses a type of graph walking algorithm called Personalised PageRank to find
    facts that are \emph{close to} the topic in the graph. It also applies
    score penalties to older facts so that recent information is preferred.

    \item \textbf{Conflict Resolution (Module C)} --- Sometimes the graph
    contains two facts that contradict each other. The system detects this,
    throws away the older, superseded fact, and checks whether the remaining
    facts actually help the LLM give a better answer. It measures this using
    entropy---if giving the LLM a fact makes it less confused, the fact is
    judged useful. A final confidence score is printed alongside the answer.
\end{enumerate}

\subsection{Important Features of the Application}

\begin{itemize}
    \item \textbf{No configuration needed} --- You literally just give it
    documents and ask questions. The schema figures itself out.
    \item \textbf{Time-aware} --- Knowing that a 2023 fact is more reliable
    than a 1990 fact is built into the scoring math.
    \item \textbf{Honest} --- If the system is not sure about an answer, it
    says so. A confidence score of 42\% means ``I found an answer in your
    documents, but my AI model is not fully convinced.''
    \item \textbf{Explainable} --- When the system removes a conflicting fact,
    it tells you: ``I discarded this because it was superseded in 2024 by a
    newer fact.''
    \item \textbf{Multi-domain} --- The same code, zero changes, works on medical
    documents, legal documents, space science, and politics.
\end{itemize}

% ----------------------------------------------------------------
\section{Problem Statement}

\subsection{Rationale for Research}

Large language models---the kind that power ChatGPT, Gemini, and similar
tools---have a frozen memory. Everything they know was learnt during training.
If something changed after that training cutoff, the model does not know. Worse,
it will often answer confidently as if it does know, which is dangerous.

Standard Retrieval-Augmented Generation (RAG) was designed to fix this. You
give the LLM your own documents at question time. But here is the catch: if your
document collection contains two articles that say different things (which is
very common in news archives, medical research, and legal databases), standard
RAG hands both conflicting pieces of text to the LLM and hopes for the best.
The LLM has no principled way to decide which one to trust.

This project was motivated by one observation: \emph{a well-designed system
should be able to look at two conflicting facts, compare their dates, count
their sources, and make a reasoned decision about which to trust}---exactly the
way a trained researcher or lawyer would.

\subsection{Barriers to Effective Research}

\begin{itemize}
    \item No open-source RAG framework detects conflicts between retrieved
    documents. LangChain, LlamaIndex---none of them do this.
    \item Knowledge-graph RAG tools require the developer to manually type out all
    the entity categories before the system will work. If you switch from politics
    to medicine, you have to rewrite configuration files.
    \item None of these systems report any confidence level. Every answer is
    presented at maximum confidence---which is simply not honest.
    \item When contradictions are resolved internally, the user is never told.
    There is no audit trail.
\end{itemize}

\subsection{Challenges in Modern Research Workflows}

The documents that researchers, students, and professionals work with tend to
have three characteristics that trip up standard RAG:

\begin{enumerate}
    \item \textbf{Facts change over time.} The chairman of ISRO in 1975 is not
    the chairman today. The first-line drug for diabetes in 1960 is not the
    recommended drug today. A system that does not track time will give wrong
    answers to questions about current state.

    \item \textbf{Important facts appear in multiple places.} If the same fact
    (e.g., ``Chandrayaan-3 landed on August 23, 2023'') appears in five documents,
    that is stronger evidence than if it appeared in only one. The system should
    reward multiply-attested facts.

    \item \textbf{Domains change completely.} A system designed for one topic
    should not break when pointed at a completely different topic. Needing to
    rewrite config files every time defeats the purpose of an intelligent system.
\end{enumerate}

\subsection{Project Objectives}

\begin{enumerate}
    \item Re-implement the TruthfulRAG v4 pipeline (arXiv:2511.10375) using
    Python, LangChain, Ollama, and Neo4j.
    \item Extend it with six improvements addressing the shortcomings listed above.
    \item Remove the one remaining manual step (schema configuration) by adding
    automatic schema inference.
    \item Test the system on multiple real-world domains---including Indian
    science, law, and medicine---to prove domain-agnosticism.
    \item Evaluate quantitatively against the v4 baseline using retrieval scores
    and conflict-detection rates.
\end{enumerate}

% ----------------------------------------------------------------
\section{Motivation}

\subsection{Purpose}

The motivation behind this project is straightforward: an AI system that can
say ``I am only 42\% confident in this answer'' is more useful for serious work
than one that always sounds 100\% confident. Real research, real legal work, and
real medical decisions require knowing when to trust a source and when to be
sceptical.

This project demonstrates that building a \emph{trustworthy} RAG system is not
a research-only endeavour. It can be built by a student, from open-source tools,
without a supercomputer, and without sending anyone's data to an external server.
If it can be done here, it can be deployed in hospitals, courts, and government
offices that cannot afford commercial AI subscriptions.
