
% ================================================================
% CHAPTER 2: LITERATURE SURVEY
% ================================================================

\section{Existing Solution Introduction}

Retrieval-Augmented Generation was formalised by Lewis et al.\ \cite{lewis2020rag}
as a method to overcome the static knowledge limitations of parametric LLMs. The
original implementation combined a Dense Passage Retriever with a BART generative
model, demonstrating measurable improvements on knowledge-intensive NLP tasks
such as open-domain question answering and fact verification.

LangChain \cite{chase2022langchain}, launched in 2022, standardised RAG into a
composable Python toolkit. Its \texttt{RetrievalQA} chain encapsulates the full
pipeline: document loading, text splitting, embedding, vector store indexing,
similarity retrieval, and prompted answer generation. This abstraction enabled
rapid prototyping but fixed a flat-retrieval architecture with no structural
knowledge representation or conflict awareness.

% ----------------------------------------------------------------
\section{Relevant Theories and Models}

\subsection{Knowledge Graphs for QA}

Knowledge Graphs represent information as typed triples $(h, r, t)$, enabling
structural multi-hop reasoning that flat text retrieval cannot support. Early KG-QA
systems such as KGQA \cite{pagerank1998} traversed Freebase to answer factoid
questions. More recent work applies embedding-based models (TransE, RotatE) to
learn distributed representations of entities and relations for link prediction.

G-Retriever \cite{he2024gretriever} demonstrated that graph-based retrieval
combined with LLM generation outperforms vanilla RAG on multi-hop textual graph
QA tasks. Microsoft GraphRAG \cite{edge2024graphrag} further showed that
community-hierarchical graph construction enables high-quality global summaries
of large corpora.

\subsection{Temporal Knowledge Graphs}

Lacroix et al.\ \cite{lacroix2020temporal} extended tensor factorisation models
to temporal KGs, where each triple carries a timestamp. CronKGQA introduced a
benchmark for temporal question answering over evolving KGs. These approaches,
however, focus on embedding-based link prediction rather than retrieval
augmentation over arbitrary text.

\subsection{Entropy and Uncertainty in LLMs}

Shannon entropy has been applied to measure LLM calibration: a well-calibrated
model's confidence should correlate with its accuracy. Kadavath et al.\ (2022)
showed that LLMs can self-assess uncertainty via sampling; Farquhar et al.\
(2023) introduced semantic entropy, grouping semantically equivalent answers
before entropy computation to avoid penalising paraphrase variation.

TruthfulRAG v4 \cite{truthfulrag2024} operationalised this into a retrieval
decision rule: a path is corrective only if providing it as context reduces the
LLM's sampling entropy by at least a learned threshold $\tau$.

% ----------------------------------------------------------------
\section{Comparison of Similar Projects}

\begin{table}[H]
\centering
\caption{Comparison of knowledge-graph RAG systems}
\label{tab:lit_compare}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{Feature} & \textbf{LangChain} & \textbf{GraphRAG} & \textbf{TruthfulRAG v4} & \textbf{v5 (Ours)} \\
\hline
Knowledge Graph storage      & No      & Yes & Yes & Yes \\
Conflict detection           & No      & No  & Yes & Yes \\
Temporal decay               & No      & No  & No  & Yes \\
Cross-document corroboration & No      & No  & No  & Yes \\
Historical snapshot queries  & No      & No  & No  & Yes \\
Explanation audit trail      & No      & No  & No  & Yes \\
Calibrated confidence score  & No      & No  & No  & Yes \\
Auto schema inference        & No      & No  & No  & Yes \\
Open source / local inference& Partial & Yes & Yes & Yes \\
\hline
\end{tabular}}
\end{table}

% ----------------------------------------------------------------
\section{Identified Gaps in the Literature}

Based on the survey above, the following gaps exist in current open-source
KG-RAG research:

\begin{enumerate}
    \item \textbf{No temporal scoring}: Existing systems do not penalise older
    facts during retrieval. A 2010 fact and a 2024 fact compete equally.
    \item \textbf{No corroboration weighting}: No system rewards facts attested
    by multiple independent sources with a higher retrieval score.
    \item \textbf{No automatic schema discovery}: All KG construction pipelines
    require manual entity-type and relation-type specification.
    \item \textbf{No calibrated output confidence}: Systems always answer at
    maximum stated confidence, giving no warning when evidence quality is poor.
    \item \textbf{No structured conflict explanation}: When contradictions are
    resolved, no system communicates the resolution rationale to the user.
\end{enumerate}

TruthfulRAG v5 addresses all five gaps simultaneously.

% ----------------------------------------------------------------
\section{Conclusion of Survey}

The literature demonstrates clear progress from flat-chunk RAG (LangChain) to
graph-structured retrieval (GraphRAG, LlamaIndex Property Graph) to
entropy-guided conflict resolution (TruthfulRAG v4). However, no system in the
literature combines temporal decay, cross-document corroboration, hybrid
retrieval, historical snapshots, explanation chains, and calibrated confidence
under a single, domain-agnostic, fully open-source pipeline. TruthfulRAG v5
fills this gap and is the primary contribution of this project.
