
% ================================================================
% CHAPTER 7: CONCLUSION
% ================================================================

\section{Contributions to the Field}

This project makes the following contributions:

\begin{enumerate}
    \item \textbf{Faithful implementation of TruthfulRAG v4}: The three-module
    architecture (Graph Construction, Graph Retrieval, Conflict Resolution) of
    arXiv:2511.10375 is implemented faithfully in Python using open-source tools,
    providing a reproducible reference implementation for the academic community.

    \item \textbf{Six novel algorithmic extensions}:
    \begin{itemize}
        \item \textbf{[N1] Cross-document corroboration scoring}: Retrieval score
        reflects the number of independent sources attesting a fact, via
        logarithmic weighting stored as a \texttt{support} property on each graph edge.
        Facts agreed upon by multiple documents score proportionally higher.

        \item \textbf{[N2] Exponential temporal decay}: Facts are penalised by
        age using $e^{-\lambda t}$ with $\lambda = 0.08$, encoding an 8.7-year
        half-life for general knowledge. The decay is applied at query time so
        the graph remains revision-free.

        \item \textbf{[N3] Hybrid BM25 + Semantic retrieval (RRF)}: Keyword and
        semantic rankings are fused via Reciprocal Rank Fusion for more robust
        retrieval, replacing single-modality cosine-only search and capturing
        queries that dense embeddings alone miss.

        \item \textbf{[N4] Explanation chain answer generation}: When contradictions
        are resolved, the system produces a structured audit trail of which knowledge
        paths were kept and which were removed, rendered visibly in the web interface
        alongside every answer.

        \item \textbf{[N5] Calibrated confidence score}: A three-component
        (entropy change, corroboration support, temporal recency) score gives users
        a calibrated uncertainty estimate as a percentage rather than false certainty.

        \item \textbf{[N6] Claim Verification Mode}: Given a declarative claim,
        the system queries its knowledge graph and returns a structured verdict
        of \texttt{SUPPORTED}, \texttt{REFUTED}, or \texttt{UNCERTAIN}, together
        with a confidence score and the specific graph evidence used. This is
        not possible in standard LangChain RAG.
    \end{itemize}

    \item \textbf{Automatic schema inference [A4+]}: A pre-pipeline LLM call
    discovers the entity and relation vocabulary from the corpus itself,
    eliminating the only remaining configuration requirement of TruthfulRAG v4
    and making the system fully domain-agnostic across medical, legal, scientific,
    and historical corpora.

    \item \textbf{Production-quality live web demonstration}: A Flask API server
    paired with a custom HTML/JavaScript dual-panel interface provides a
    side-by-side comparison of standard LangChain RAG against TruthfulRAG v5 on
    any user-supplied query. The interface renders knowledge paths, confidence bars,
    conflict counters, and claim verification results in real time.

    \item \textbf{Robustness hardening}: Apostrophe injection in Neo4j Cypher
    strings, \texttt{KnowledgePath} attribute serialisation, corpus dropdown
    activation state, Windows UTF-8 encoding, and APOC availability failure
    modes are all handled gracefully, making the codebase suitable for deployment
    beyond a research prototype.
\end{enumerate}

% ----------------------------------------------------------------
\section{Future Enhancements}

\begin{itemize}
    \item \textbf{Multi-hop graph reasoning}: Current retrieval operates on
    single edges. Future work should implement path-following retrieval (BFS or
    PPR on paths of length $> 1$) to answer complex multi-relation queries such
    as ``Who founded the company that acquired Twitter?''

    \item \textbf{Persistent, accumulative knowledge base}: Rather than clearing
    and rebuilding the Neo4j graph on each corpus load, future versions should
    maintain a persistent graph where new documents add facts and contradictions
    update old facts with a validity end timestamp rather than deleting them.

    \item \textbf{LangGraph agentic loop}: Wrapping the pipeline as a LangGraph
    node would enable an agent to retry retrieval with a broadened query if
    confidence falls below a configurable threshold.

    \item \textbf{Cross-encoder re-ranking}: After PPR scoring, a cross-encoder
    model (e.g., \texttt{ms-marco-MiniLM-L-12-v2}) could be used as a final
    re-ranking step to push the most precisely relevant path to the top.

    \item \textbf{Multilingual corpus support}: Replacing
    \texttt{all-MiniLM-L6-v2} with \texttt{paraphrase-multilingual-MiniLM-L12-v2}
    and a multilingual LLM (e.g.\ Aya-23) would extend the system to Hindi,
    Tamil, Meitei, and other regional languages without any architectural change.

    \item \textbf{Cross-corpus graph persistence}: Currently, loading a new
    corpus clears the previous graph. A future version should support
    named, coexisting graph partitions, allowing simultaneous retrieval
    across medical, legal, and scientific domains in a single query.

    \item \textbf{Year-anchored temporal decay} (\emph{resolved in v5 final}):
    An early version computed decay relative to the current calendar year for
    all queries, which incorrectly penalised facts from the target year when
    the user asked a year-specific question (e.g.\ ``Who was the CEO in 2020?'').
    The final implementation detects the target year from the query text and
    anchors the decay reference to that year, so facts from exactly the target
    year receive full weight ($w = 1.0$) rather than being discarded as stale.
\end{itemize}

% ----------------------------------------------------------------
\section{Final Thoughts}

TruthfulRAG v5 demonstrates that academically rigorous, conflict-aware RAG is
achievable using entirely open-source components on consumer hardware. The
combination of structured knowledge representation (Neo4j), entropy-guided
conflict resolution (adapted from arXiv:2511.10375), and calibrated uncertainty
quantification produces a system that is not only more accurate than standard
RAG on temporally-conflicted corpora, but also more transparent and honest
about the limits of its own knowledge.

The live demonstration interface makes this difference concrete and visible.
When both systems receive the same query against a corpus containing a temporal
conflict, standard LangChain RAG returns an answer with no qualification.
TruthfulRAG v5 returns the same factual content, adds a confidence percentage,
shows which facts were retained and which were filtered, and offers a claim
verification endpoint that returns a structured verdict rather than free-form
prose. These are not cosmetic differences; they represent a fundamental shift
in what a retrieval-augmented system is accountable for.

The most important design insight of this project is that \emph{honesty is a
feature, not a limitation}. A system that reports ``42\,\% confidence'' when its
internal certainty is low is more valuable in a high-stakes setting than one
that confidently outputs a stale or incorrect answer. TruthfulRAG v5 is a
step towards AI systems that know when they do not know.


